---
title: "Play with Data v0.1"
author: "Constantinos Yeles"
date: "April 23, 2019"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
params:
  date: !r lubridate::today()
editor_options:
  chunk_output_type: inline
---

#### How to make, move & copy folders
While working with data and analyzing them we need to organize our results in specific folders.   
Fortunately we could perform folder / file manipulation using R.   
```{r Folder handling}
# 1.1 make a folder                                                         ----
help(dir.create)#or ?dir.create()  get help about this function 
dir.create("my_Data")

# we can give names to a vector and use it to make more folders
fold_names <- c("my_data", "my_plots", "my_results")
dir.create(fold_names) # we get an Error

# we need to use for or sapply to create multiple folders
sapply()
sapply(fold_names, function(x) dir.create(x)) 
# or with for (1):
for (i in seq(along=fold_names)) {
  folder <- fold_names[i]
  dir.create(folder)
}
# with for (2):
for (i in fold_names) {
    dir.create(i)
}

# if we have different datasets it would be more convinient to 
# have different folders for each dataset
dat_names <- c("my_1st_dataset", "my_2nd_dataset") 
dat_names # or we could use glue package
name_of_datasets <- c("methylation", "smallRNA", "exome" )
dat_names <- str_glue("my_{name_of_datasets}_dataset") 
dat_names
# make folders
sapply(dat_names, dir.create) 
# 1.2 copy or move folders                                                  ----
# we want to have in each dataset folder various subfolders:
# "my_data", "my_plots" and "my_results"
# for one folder:
file.copy(fold_names,"my_methylation_dataset",recursive = TRUE)
# for all:
sapply(dat_names, function(x) file.copy(fold_names,x,recursive = TRUE))

# to move a folder we could use rename which is used also for renaming folder/files
file.rename("my_Data", "my_exome_dataset/my_Data_1")
```
#### Working directory
A working directory in R is the main folder, the current working environment,  
in which all user-defined objects are contained. 
```{r working dirs}
# get the working directory
getwd()
# set the working directory
setwd("my_smallRNA_dataset/")
```
### Basic manipulation of data
#### Example Dataset
Let's make a gene set data set 
```{r "Dataset"}
# I always prefer to tag my results with the date of the analysis
# in the name of resulted files. so we can use :
Sys.time()
# I prefer to use only the date with the name of the month and underscore "_":
date <- gsub(" ","_",format(Sys.time(), "%b %d %Y"))
date
# load packages
library(tidyverse)
library(annotables)# to install run : BiocManager::install( "stephenturner/annotables")
set.seed(313) # reproducibility

FileName <- rep(paste0("file", 1:5,".txt"),each=10) # character vector with
#different file names of our aligned samples
FileName
# Same output different function (str_glue)
number_of_files <- c(1:5) # two sample files, each for 1 condition
FileName <- rep(str_glue("file{number_of_files}.txt"),each = 10) %>% as_factor()
FileName

read_id <-  rep(paste0("t00", 1:4,1000:1009), 5) %>% as_factor() # factor with
# id of reads "aligned"
read_count <-  sample(1:10000, 50)  # integer with number of reads  
ENST <- rep(sample(grch38_tx2gene$enstxp,10),5) %>% as_factor() # factor with
# transcript id

# use sapply to create sampl read sequences
read_sequence <- rep(sapply(c(31:40),
          function(x) 
          paste(sample(c("A","C","T","G"),x, replace=TRUE), collapse="")),5)

# create the dataset in long format using tibble function
test_data <- tibble::tibble(FileName,read_id,read_count,ENST,read_sequence)
test_data

# if we want to check how many reads are per "library" then we could use
# summarise function
summarise(test_data,sum(read_count))# not very informative
# we want per library, so we need to group first by "library" our dataset and then use summarise.
test_data %>% # using pipe %>% for easier implementation, help(%>%) to learn more # help(%>%) to learn more the "pipe"
  group_by(FileName) %>% # we group it by each "FileName" which equals to a sample library
  summarise(sum(read_count))

# tables are good for interpretation of the data but we could plot them easily
# with ggplot
test_data %>% 
  group_by(FileName) %>% 
  summarise( LibSum = sum(read_count)) %>% 
  ggplot() +
  geom_col(mapping = aes(x=FileName, y=LibSum, fill = FileName))

```
why [set.seed](https://stackoverflow.com/questions/13605271/reasons-for-using-the-set-seed-function)?
